Historical trends in the contribution of six major causes of death to the black-white

### 2017-04-28 Manuscript/Manuscript_COD_Trends.Rmd
added infant contrib plots

```{r load_data, echo=F}
BlackWhite_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results2/BlackWhite_results.csv")
BlackWhite_results <- reorder.as.map(BlackWhite_results, "state", "stabbrs")
cod_decomp_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results2/cod_decomp_results.csv")
cod_decomp_results <- reorder.as.map(cod_decomp_results, "state", "stabbrs")
cod_decomp_results$COD_rev <- factor(cod_decomp_results$COD, levels = c("Cardiovascular", "Cancers", "All other causes",
                                                                        "Communicable", "Non-communicable",  "Injuries"))
```

@@

  c <- ggplot(subset.data,
              aes(x = year, y = COD_cont_yrs_mean)) + 
    geom_hline(aes(yintercept = 0)) + geom_vline(aes(xintercept = 1969)) + #geom_hline(yintercept = 0, col = "black") +
    geom_line(aes(y = fitted.RE - RE.estimates), col = "#d7191c") +
    geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl), fill = "#abd9e9", alpha = 0.65) +
    geom_line(col = "#08519c") + 
    #geom_line(aes(col = Census_Division)) + #add this for the online version, and then adjust legend and hover display    
    #scale_fill_manual(values = c("#b2e2e2")) +
    facet_wrap( ~ stabbrs.map.order, ncol = 11, drop = F) + 
    #geom_line(aes(y = fitted.RE), col = "#d7191c", lty = 3) +
    ggtitle(paste0("Estimated smoothed contribution of ", COD1, " disease to the black-white life expectancy gap (blue 95% credible interval) vs. the meta-estimated general pattern (red line) in ", sex1, " 1969-2013")) +
    theme_classic(base_size = 10) + 
    theme(axis.text.x = element_blank(),
          strip.background=element_blank(),
          axis.line=element_blank(),
          axis.ticks=element_blank()) +
    ylab("Contribution to the life expectancy gap (years)") +
    xlab("Year (1969-2013)")

@@

@@

@@

>>>>

```{r load_data, echo=F}
BlackWhite_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results2/BlackWhite_results.csv")
BlackWhite_results <- reorder.as.map(BlackWhite_results, "state", "stabbrs")
cod_decomp_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results2/cod_decomp_results.csv")
cod_decomp_results <- reorder.as.map(cod_decomp_results, "state", "stabbrs")
cod_decomp_results$COD_rev <- factor(cod_decomp_results$COD, levels = c("Cardiovascular", "Cancers", "All other causes",
                                                                        "Communicable", "Non-communicable",  "Injuries"))

age_cod_results_infants_AOC <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results2/age_cod_results_infants_AOC.csv")
age_cod_results_infants_AOC$stabbrs <- state.abb[match(age_cod_results_infants_AOC$state, state.name)]
age_cod_results_infants_AOC$stabbrs[age_cod_results_infants_AOC$state == "Washington DC"] <- "DC"
age_cod_results_infants_AOC <- reorder.as.map(age_cod_results_infants_AOC, "state", "stabbrs")

@@

  c <- ggplot(subset.data,
              aes(x = year, y = COD_cont_yrs_mean)) + 
    geom_hline(aes(yintercept = 0)) + geom_vline(aes(xintercept = 1969)) + #geom_hline(yintercept = 0, col = "black") +
    geom_line(aes(y = fitted.RE - RE.estimates), col = "#d7191c") +
    geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl), fill = "#abd9e9", alpha = 0.65) +
    geom_line(col = "#08519c") + 
    #geom_line(aes(col = Census_Division)) + #add this for the online version, and then adjust legend and hover display    
    #scale_fill_manual(values = c("#b2e2e2")) +
    facet_wrap( ~ stabbrs.map.order, ncol = 11, drop = F) + 
    #geom_line(aes(y = fitted.RE), col = "#d7191c", lty = 3) +
#    ggtitle(paste0("Estimated smoothed contribution of ", COD1, " disease to the black-white life expectancy gap (blue 95% credible interval) vs. the meta-estimated general pattern (red line) in ", sex1, " 1969-2013")) +
    theme_classic(base_size = 10) + 
    theme(axis.text.x = element_blank(),
          strip.background=element_blank(),
          axis.line=element_blank(),
          axis.ticks=element_blank()) +
    ylab("Contribution to the life expectancy gap (years)") +
    xlab("Year (1969-2013)")

@@

```{r, echo = F}
male.endpoints.aoc <- cod_decomp_results %>% filter(year %in% c(1969, 2013), sex =="Male", COD == "All other causes", state == "Alabama") %>% transmute(round(fitted.RE - RE.estimates, 2) )

female.endpoints.aoc <- cod_decomp_results %>% filter(year %in% c(1969, 2013), sex =="Female", COD == "All other causes", state == "Alabama") %>% transmute(round(fitted.RE - RE.estimates, 2) )

male.infants.endpoints.aoc <- age_cod_results_infants_AOC %>% filter(year %in% c(1969, 2013), sex =="Male", state == "Alabama") %>% transmute(round(fitted.RE - RE.estimates, 2) )

female.infants.endpoints.aoc <- age_cod_results_infants_AOC %>% filter(year %in% c(1969, 2013), sex =="Female", state == "Alabama") %>% transmute(round(fitted.RE - RE.estimates, 2) )
```

@@

*CR note: I just added the following paragraph. Open to how/whether we want to talk about this. Thought we need to say something about how "all other causes partiallly reflects infant mortality. Also these additional figures (Fig 15 and 16) would be moved to the appendix.*

To investigate how much of the gap in all other causes occurred during the first year of life, we conducted a sensitivity analysis to calculate the contribution of deaths due to all other causes that occurred *during the first year of life* to the life expectancy gap (**Figures 15 and 16**). We investigated this because the majority of infant deaths are classified as being due to all other causes, conflating the all other causes category with inquities in infant mortality. We found that 43% of the contribution of all other causes to the gap was rooted in differences during the first year of life. In 2013, 53% of these deaths occurred during the first year of life. Thus, the trend in the all other causes contribution partially reflects a decreasing disparity in infant mortality.

@@

**Figure 15: Smoothed state-level trends in the contribution of all other causes to the life expectancy gap vs. the overall pattern in males younger than one**

```{r age-cod-trend-infant-males, fig.width = 9, fig.height = 9, echo = F}
ggplot(dat = subset(age_cod_results_infants_AOC, sex == "Male"), aes(y = age_COD_cont_yrs_mean, x = year)) +
  geom_line(aes(y = fitted.RE - RE.estimates), col = "#d7191c") +
  geom_hline(aes(yintercept = 0)) + geom_vline(aes(xintercept = 1969)) +
  geom_ribbon(aes(ymin = age_COD_cont_yrs_lcl, ymax = age_COD_cont_yrs_ucl), fill = "#abd9e9", alpha = 0.65) +
  geom_line(col = "#08519c") +  
  facet_wrap( ~ stabbrs.map.order, ncol = 11, drop = F) +
  theme_classic(base_size = 10) + 
  theme(axis.text.x = element_blank(),
        strip.background=element_blank(),
        axis.line=element_blank(),
        axis.ticks=element_blank()) +
  ylab("Contribution to the life expectancy gap (years)") +
  xlab("Year (1969-2013)") +
  geom_hline(aes(yintercept = 0), lwd = 0.5, col = "grey", alpha = 0.5) + 
  geom_hline(aes(yintercept = 1), lwd = 0.5, col = "grey", alpha = 0.5) +
  geom_hline(aes(yintercept = 2), lwd = 0.5, col = "grey", alpha = 0.5)
```

**Figure 16: Smoothed state-level trends in the contribution of all other causes to the life expectancy gap vs. the overall pattern in females younger than one**

```{r age-cod-trend-infant-females, fig.width = 9, fig.height = 9, echo = F}
ggplot(dat = subset(age_cod_results_infants_AOC, sex == "Female"), aes(y = age_COD_cont_yrs_mean, x = year)) +
  geom_line(aes(y = fitted.RE - RE.estimates), col = "#d7191c") +
  geom_hline(aes(yintercept = 0)) + geom_vline(aes(xintercept = 1969)) +
  geom_ribbon(aes(ymin = age_COD_cont_yrs_lcl, ymax = age_COD_cont_yrs_ucl), fill = "#abd9e9", alpha = 0.65) +
  geom_line(col = "#08519c") +  
  facet_wrap( ~ stabbrs.map.order, ncol = 11, drop = F) +
  theme_classic(base_size = 10) + 
  theme(axis.text.x = element_blank(),
        strip.background=element_blank(),
        axis.line=element_blank(),
        axis.ticks=element_blank()) +
  ylab("Contribution to the life expectancy gap (years)") +
  xlab("Year (1969-2013)") +
  geom_hline(aes(yintercept = 0), lwd = 0.5, col = "grey", alpha = 0.5) + 
  geom_hline(aes(yintercept = 1), lwd = 0.5, col = "grey", alpha = 0.5) +
  geom_hline(aes(yintercept = 2), lwd = 0.5, col = "grey", alpha = 0.5)
```
